[Re] Chaos in a long-term experiment with a plankton community

### 2015-09-25 article/article.tex
updates

The raw data show populations dynamics at least very similar to those in
figure 1b-g of the original publication (figure \ref{fig:dynamics}).

\begin{figure}[htbp]
\centering
\includegraphics{figures/unnamed-chunk-21-1}
\caption{\label{fig:dynamics}Observed population dynamics.}
\end{figure}

@@

Correlations among species abundances presented in Table 1 of the
original article closely matched our reproduced correlations, calculated
from the transformed data with zeros removed (figure
\ref{fig:corr_comp}). Deviations between the original and reproduced
correlations are relatively small (less than 0.072 units) and
infrequent. \textbf{These may have resulted from us removing zeros after
interpolation. Maybe we should check this.}

@@

@@

@@

@@

@@

Only done for non-linear models (the original article also does this for
linear models).

@@

Conclusion, at the very minimum, should indicate very clearly if you
were able to replicate original results. If it was not possible but you
found the reason why (error in the original results), you should exlain
it.

\begin{longtable}[c]{@{}llllll@{}}
\caption{\label{tbl:table}Table caption }\tabularnewline
\toprule
Heading 1 & & & Heading 2 & &\tabularnewline
\midrule
\endfirsthead
\toprule
Heading 1 & & & Heading 2 & &\tabularnewline
\midrule
\endhead
cell1 row1 & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 &
cell6 row 1\tabularnewline
cell1 row2 & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 &
cell6 row 2\tabularnewline
cell1 row3 & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 &
cell6 row 3\tabularnewline
\bottomrule
\end{longtable}

A reference to table \ref{tbl:table}. A reference to figure
\ref{fig:logo}. A reference to equation \ref{eq:1}. A reference to
citation \textcite{markdown}.

\begin{figure}[htbp]
\centering
\includegraphics{rescience-logo.pdf}
\caption{\label{fig:logo}Figure caption}
\end{figure}

\begin{equation} A = \sqrt{\frac{B}{C}} \label{eq:1}\end{equation}

>>>>

The reproduced populations dynamics were at least very similar to those
in figure 1b-g of the original publication (figure \ref{fig:dynamics}).
Note that we plot fourth root transformed values, rather than raw
abundances with a y-axis break, as in the original article.

\begin{figure}[htbp]
\centering
\includegraphics{figures/dynamics.pdf}
\caption{\label{fig:dynamics}Observed population dynamics.}
\end{figure}

@@

Correlations among species abundances presented in Table 1 of the
original article closely matched our reproduced correlations, calculated
from the transformed data with zeros removed (figure
\ref{fig:corr_comp}). Deviations between the original and reproduced
correlations are relatively small (less than 0.072 units) and
infrequent.

@@

Highlighted in the text of the original paper were: negative
correlations of picophytoplankton with protozoa, and of
nanophytoplankton both with rotifers and calanoid copepods, positive
correlation of picophytoplankton with calanoid copepods, negative
correlation between bacteria and ostracods, and positive correlation
between bacteria and phosphorus. All of these correlations were
reproduced.

@@

Spectral analyses in the original paper were presented graphically in
figures S3 (raw spectrograms) and S4 (Welch periodograms). These graphs
were, apparently, visually inspected to reveal: ``fluctuations covered a
range of different periodicities'', and ``picophytoplankton, rotifers
and calanoid copepods seemed to fluctuate predominantly with a
periodicity of about 30 days.'' It is unclear how these conclusions were
derived from figures S3 and S4 of the original article. Our reproduced
spectra (not shown here, but code provided) were not quantitatively
identical to the spectra in the original article.

@@

Reproduced divergence rates (figure \ref{fig:divergence}) and comparison
of the original and reproduced Lyapunov exponents (figure
\ref{fig:LE_comparison}). The original article states: ``the distance
between initially nearby trajectories increased over time, and reached a
plateau after about 20--30 days''. The reproduced results appear not
inconsistent with this statement, except for one group of species
(Harpacticoids). The original article also stated that the analyses
``yielded significantly positive Lyapunov exponents of strikingly
similar value for all species (Fig. 3; mean exponent = 0.057 per day,
s.d. = 0.005 per day, n = 9)''. Reproduced exponents had very similar
mean value, but had about four times greater standard deviation (mean =
0.055 and s.d. = 0.019).

\begin{figure}[htbp]
\centering
\includegraphics{figures/unnamed-chunk-50-1.pdf}
\caption{\label{fig:divergence}Reproduced divergence rates and Lyapunov
exponents (figure 3 in the original article).}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/LE_comparison.pdf}
\caption{\label{fig:LE_comparison}Comparison of Lyapunov exponents in
the original article and this reproduction.}
\end{figure}

@@

The original paper reported global Lyapunov exponent calculated via two
modelling approaches (neural network and generalised additive models
{[}GAMs{]}). Only the GAM approach was reproduced, with the assistance
of code donated by Stephen Ellner. The original article obtained a
global Lyapunov exponent of λ=0.08 day-1. The reproduced value was 0.04.
We did not reproduce the bootstrapping used to give confidence intervals
around this estimate.

@@

The article stated: ``For short-term forecasts of only a few days, most
species had a high predictability of R2 = 0.70 -- 0.90 (Fig. 2).
However, the predictability of the species was much reduced when
prediction times were extended to 15--30days.'' The reproduced
predictabilities, which were calculated from the GAMs, were consistent
with these qualitative statements, though were quantitatively different
(\ref{fig:prediction_distance}). We did not reproduce the predictability
estimates for linear models.

\begin{figure}[htbp]
\centering
\includegraphics{figures/prediction_distance.pdf}
\caption{\label{fig:prediction_distance}Predictability (correlation
between predicted and observed abundances) and prediction distance
(days) (figure 2 in the original article).}
\end{figure}

@@

Although we were not able to make a quantitatively accurate reproduction
of all results of the original article, the qualitative results were
largely identical. Quantitative differences may have resulted from
difference in algorithms used. For example,the original used the
\href{http://www.mpipks-dresden.mpg.de/~tisean/}{Tisean software} to
calculate Lyapunov exponents. As this was available from CRAN
\href{http://cran.r-project.org/web/packages/RTisean/index.html}{until
mid 2014} and since it is a bit less well integrated with R, we instead
use the tseriesChaos package \textcite{tseriesChaos}, which in any case
was largely inspired by the TISEAN project. There may also have been
some difference in data used for specific analyses, e.g., data with
zeros removed or not, as it was not always possible to be sure the
reproduction used exactly the same data.
